---
title: "ESm 244 Lab 4"
author: "Alyssa Winchell"
date: "January 31, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Attach packages

```{r include = FALSE}

library(tidyverse)
library(Ecdat)
library(plm)
library(lmtest)
library(car) #for plotting

```

1. Part 1: Nonlinear least squares (logistic growth of bacteria)

```{r}

CellGrowth <- read_csv("CellGrowth.csv")

#Look at it
ggplot(CellGrowth, aes(x = Time, y = CellCount)) + 
  geom_point()

#find estimates for log growth for training model

# N = A/(I + Be^(-rt)) r helps us find A, b, and r

# estimate for carrying capacity: 3700
# starting pop value (N0) is 2000, what is B? ~0.85
# r ~ 0.1035

BEstimate <- (3700-2000)/2000
BEstimate

#pop is growing exp from about time 0 - 5. model as alog for a linear relationship. this estimates r for growth rate coef
lag_phase <- CellGrowth %>%
  filter(Time < 5)

#plot cell growth vs time to make sure its linear
ggplot(lag_phase, aes(x = Time, y = log(CellCount)))+
  geom_point() #looks close to linear, which means that part was def exp

lm(log(CellCount) ~ Time, data = lag_phase) # r ~ 0.1035

```

Find model A, B, and r using Gauss-Newton method for iterative convergence (nonlinear least squares):

Use the nls() fxn:

```{r}

cell_fit <- nls(CellCount ~ A/(1 + B*exp(-r*Time)),
                start = list(A = 3700, B = 0.85, r = 0.1035),
                             data = CellGrowth,
                             trace = TRUE)

cell_fit

#store coeffs
A <- coef(cell_fit)[1]
B <- coef(cell_fit)[2]
r <- coef(cell_fit)[3]

#create a new seq of values that we want to make predictions for with model. New seq of time values of relevant times (time_seq)
time_seq <- seq(1,20,length = 100)

#plug new seq into model with the parameters A, B , and r that we found:
cell_pred <- A/(1 + B*exp(-r*time_seq))

#bind that together with original data (time_seq data):
pred_df <- data.frame(time_seq, cell_pred)

#create graph with original data and our model preds:
ggplot(CellGrowth, aes(x = Time, y = CellCount)) +
  geom_point(color = "Blue", size = 3) +
  geom_line(data = pred_df, aes(x = time_seq, y = cell_pred), colour = "orange", size = 1)

```

Panel Regression - First Shot

What is the effect of price on purchases for cigarettes?

entity: State
time variable: Year
packpc: our dv, pack per capita
avgprs: predictor variable of interest, average price

```{r}

cigs_panel <- Cigarette %>%
  select(state, year, packpc, avgprs)

ggplot(cigs_panel, aes(x = avgprs, y = packpc, group = state)) +
  geom_point(aes(color = state)) +
  geom_smooth(method = "lm", aes(color = state), se = FALSE, size = 0.2) #doesnt include stand eror

ggplot(cigs_panel, aes(x = year, y = packpc, group = state)) +
  geom_point(aes(color = state)) +
  geom_smooth(method = "lm", aes(color = state), se = FALSE, size = 0.2)

```

Entity fixed effects model:

```{r}

cigs_model <- plm(packpc ~ avgprs,
                  data = cigs_panel,
                  index = c("state", "year"),
                  model = "within") #EFE model
cigs_model
#if done with lm, would have 50 different levels for state = 49 non reference states

```

When we account for entity fixed effects, we would expect to see a decrease of -0.33 packs per person for every unit decrease in price ($1).

Robust Stand error for panel regression on large data sets:
```{r}

coeftest(cigs_model, vcov. = vcovHC(cigs_model, type = "HC1")) #more conservatice estimate of relationship, esp with regards to stand error = slightly different coef.

```

Other things are likely affecting cig trends in all states. Add time fixed effects:

```{r}

et_cigs_model <- plm(packpc ~ avgprs,
                     data = cigs_panel,
                     index = c("state", "year"),
                     model = "within",
                     effect = "twoways")
et_cigs_model

#in console: ?pFtest. is there a differnce between the models? is the model improved? 
pFtest(et_cigs_model, cigs_model)

```

There is a sig additive effect of adding a time fixed effect.
